\chapter*{Abstract}

This project addresses a critical challenge in gaming accessibility through the development of a sophisticated Configuration Graphical User Interface (GUI) for MotionInput, an adaptive software system enabling alternative computer interaction methods. The project transforms the complex, technical process of JSON-based configuration into an intuitive, visual interface, significantly reducing barriers for users with physical disabilities.

The implementation leverages modern technologies including the WinUI 3 framework and Stable Diffusion AI, architected using MVVM patterns and service-oriented design principles. The GUI integrates seamlessly with the MotionInput ecosystem while introducing innovative features such as AI-powered icon generation, real-time configuration validation, and an intuitive profile management system.

Key features include a visual profile editor with real-time preview capabilities, an AI-assisted icon studio, comprehensive action configuration tools, and a robust profile management system. The implementation adheres to WCAG 2.1 accessibility guidelines and incorporates extensive user feedback, supporting both basic and advanced configuration scenarios while maintaining high performance and reliability.

Evaluation results demonstrate significant improvements in accessibility and user experience, with user satisfaction rates reaching 90\% in testing. The project establishes a new paradigm for accessible gaming configuration, combining technical sophistication with user-centric design to enhance gaming accessibility for users with diverse physical abilities.

\clearpage
